---
title: "Ask Abby Analysis"
output:
  html_document:
    theme: readable
    df_print: paged
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hold', warning = FALSE, fig.width = 12)
source("clean.R")
source("theme_lanyop.R")
```

# Check Distributions

### Numerics:

We want to see the distribution of each of our numeric variables.  We'll be handling these like numeric categoricals.
```{r num-dist, results='hide', warning=FALSE}
# Identify numeric cols
num_vars <- da_data %>% select_if(is.numeric) %>% names

# Build plots in loop
plots <- da_data %>% 
  map_if(is.numeric, function(var) dist_plot(da_data, var)) %>% 
  .[num_vars]
```

Now we can print all of them at once.
```{r num-plots, results='hold'}
# Print resulting plots
print(
  map(seq_along(plots), 
      function(i) {
        plots[[i]] + 
        labs(title = paste0("Distribution of ", names(plots[i])),
             x = names(plots[i]))
      }))
```

### Text:

The text will need to be handled like natural language.  We'll be use `tidytext` to drill into the question data.

#### Lemma Analysis:

Let's dig into the lemmas (related word groupings) for our text by decade and year.
```{r txt-lemmas}
# Bind tf-idf metric
metric_data <- lemma_data %>% 
  count(id, lemmas) %>% 
  bind_tf_idf(term = "lemmas", document = "id", n = "n" ) %>% 
  group_by(lemmas) %>% 
  summarize(tf_idf = mean(tf_idf),
            total  = sum(n)) %>% 
  ungroup 

# Use tf-idf to get common lemmas
common_lemmas <- metric_data %>% 
  filter(tf_idf < 0.5) %>% 
  arrange(-total) %>% 
  pull(lemmas)

# Count all lemmas over time
lemmas_by_time <- 
  lemma_data %>% 
  group_by(year, decade, lemmas) %>% 
  summarize(year_counts = n()) %>% 
  ungroup %>% 
  group_by(decade, lemmas) %>% 
  mutate(decade_counts = sum(year_counts)) %>% 
  ungroup %>% 
  arrange(year, -year_counts)
```

Let's see which lemmas are the most used in letters in each decade.
```{r txt-}
# Get decade only counts 
lemmas_by_decade <- lemmas_by_time %>% 
  select(decade, lemmas, decade_counts) %>% 
  distinct %>% 
  arrange(decade, -decade_counts)

# Plot top 10 by decade
lemmas_by_decade %>% 
  group_by(decade) %>% 
  slice(1:10) %>% 
  ungroup %>% 
  mutate(decade = as_factor(as.character(decade)),
         lemmas = as_factor(lemmas) %>% fct_rev) %>% 
  ggplot(aes(x = lemmas, y = decade_counts)) +
  geom_col(aes(fill = decade), show.legend = F) +
  scale_fill_manual(values = lanyop_pal("all")(4)) +
  coord_flip() +
  facet_wrap(facets = "decade", scales = "free") +
  labs(title = "Top 10 Lemmas By Decade") +
  theme_lanyop()
```

Now let's use the `common_lemmas` we found to filter down to a more unique top ten for each decade.
```{r}
# Top 10 by decade (filtered)
lemmas_by_decade %>% 
  filter(!lemmas %in% common_lemmas[1:100]) %>%
  arrange(decade, -decade_counts) %>% 
  group_by(decade) %>% 
  slice(1:10) %>% 
  ungroup %>% 
  mutate(decade = as_factor(as.character(decade)),
         lemmas = as_factor(lemmas) %>% fct_rev) %>% 
  ggplot(aes(x = lemmas, y = decade_counts)) +
  geom_col(aes(fill = decade), show.legend = F) +
  scale_fill_manual(values = lanyop_pal("all")(4)) +
  coord_flip() +
  facet_wrap(facets = "decade", scales = "free") +
  labs(title = "Top 10 Lemmas By Decade (filtered for uniqueness)", 
       subtitle = "Group avg tf-idf > 0.5") +
  theme_lanyop()
```

Now let's see how lemmas shared by all four decades usage changes over time.
```{r}
# Find top 8 lemmas from the overall set
top_lemmas <- lemmas_by_decade %>% 
  filter(!lemmas %in% common_lemmas[1:100]) %>% 
  group_by(lemmas) %>% 
  summarize(total = sum(decade_counts)) %>% 
  ungroup %>% 
  arrange(-total) %>% 
  slice(1:8) %>% 
  pull(lemmas)

# Plot percent of total uses by year
lemmas_by_time %>% 
  filter(lemmas %in% top_lemmas) %>% 
  group_by(lemmas) %>%
  mutate(ovr_counts = sum(year_counts)) %>% 
  ungroup %>% 
  ggplot(aes(year, year_counts / ovr_counts)) +
  geom_point() +
  geom_smooth(method = "loess") +
  facet_wrap(~ lemmas, scales = "free_y", ncol = 4) +
  scale_y_continuous(labels = scales::percent_format()) +
  ylab("% frequency of lemma by year") +
  theme_lanyop()
```

The usage of these share a similar pattern in most cases, where there's a dip in usage in 1990's.  Are there just less letters from then?
```{r}
# Plot percent of letters by year
letters_per_year <- 
  da_data %>% 
  count(year) %>% 
  mutate(total_letters = sum(n),
         percent_total = n / total_letters)

letters_per_year %>% 
  ggplot(aes(year, percent_total)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Percent of Total Letters Contributed By Year") +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_lanyop()
```

It appears that there are indeed less letters from the mid 1990's!  So then, which of those top 6 lemmas differ significantly from this pattern?
```{r}
# Find percent of letters containing word relative to number of letters in each time period
lemmas_by_time %>% 
  filter(lemmas %in% top_lemmas) %>%
  # group_by(lemmas) %>%
  # mutate(ovr_counts = sum(year_counts)) %>% 
  # ungroup %>% 
  left_join(letters_per_year, by = "year") %>% 
  mutate(rel_freq = year_counts / total_letters)
  
  ggplot(aes(year, rel_freq)) +
  geom_point() +
  geom_smooth(method = "loess") +
  facet_wrap(~ lemmas, scales = "free_y", ncol = 4) +
  scale_y_continuous(labels = scales::percent_format()) +
  ylab("% frequency of lemma by year") +
  theme_lanyop()
```


#### n-gram Analysis:

We should have enough data to get back interesting lemmas by year.
```{r}
# Get bigrams over time
bigrams_by_time <- 
  bigram_data %>% 
  group_by(year, decade, bigrams) %>% 
  summarize(year_counts = n()) %>% 
  ungroup %>% 
  group_by(decade, bigrams) %>% 
  mutate(decade_counts = sum(year_counts)) %>% 
  ungroup %>% 
  arrange(year, -year_counts)
```

Now let's plot the top 10 bigrams over time just like we did with lemmas.
```{r}
# Plot top 10 bigrams by decade
bigrams_by_time %>% 
  select(decade, bigrams, decade_counts) %>% 
  distinct %>% 
  arrange(decade, -decade_counts) %>% 
  group_by(decade) %>% 
  slice(1:10) %>% 
  ungroup %>% 
  mutate(decade = as_factor(as.character(decade)),
         bigrams = as_factor(bigrams) %>% fct_rev) %>% 
  ggplot(aes(x = bigrams, y = decade_counts)) +
  geom_col(aes(fill = decade), show.legend = F) +
  scale_fill_manual(values = lanyop_pal("all")(4)) +
  coord_flip() +
  facet_wrap(facets = "decade", scales = "free") +
  labs(title = "Top 10 bigrams By Decade") +
  theme_lanyop()
```
